---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# likelihoodTools: Tools for Managing Maximum Likelihood Estimation Results

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/ajpelu/likelihoodTools/graph/badge.svg)](https://app.codecov.io/gh/ajpelu/likelihoodTools)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/ajpelu/likelihoodTools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ajpelu/likelihoodTools/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The likelihoodTools package provides tools for managing and exploring parameter 
estimation results derived from Maximum Likelihood Estimation (MLE) using the 
[likelihood](https://cran.r-project.org/web/packages/likelihood/index.html) package. This package simplifies analysis workflows by offering 
functions that support the organization, visualization, and summary of MLE 
outcomes, aiding in deeper statistical insights.

The package is designed to simplify the exploration and interpretation of MLE results, 
providing a comprehensive set of tools for managing parameter estimation outcomes.

## Installation

You can install the development version of likelihoodTools from 
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ajpelu/likelihoodTools")
```

## Key Features

- Tools for organizing MLE outputs.
- Functions to explore, visualize, and interpret parameter estimation results.
- Integrations with dplyr and ggplot2 for efficient data manipulation and visualization.

## Citation

If using this package, please cite it:

```{r comment=NA}
citation("likelihoodTools")
```

## Funding 

Antonio J. Pérez-Luque received funding from the Spanish Ministry of Science and 
Innovation through the Juan de la Cierva postdoctoral fellowship program. 
Grant JDC2022-050056-I funded by MCIN/AEI/ 10.13039/501100011033 and by “European Union NextGenerationEU/PRTR


![](https://drive.google.com/file/d/1mMqNhNAeIlEyY_QTjZZ7bLtV96pg0ww-/view?usp=sharing)





