---
title: "Getting started"
editor_options: 
  chunk_output_type: console
bibliography: grateful-refs.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  fig.height = 4.5,
  fig.width = 4.5,
  fig.align = "center",
  comment = "#>"
)
```

```{r pkgs}
library(likelihood)
library(likelihoodTools)
```

## Introduction

This vignette will show you how to get started with the `likelihoodTools` package. 
The `likelihoodTools` package provides a set of tools to help you work with the 
`likelihood` package [@likelihood]. 

We use the dataset and the example included in the `likelihood` to 
illustrate how to use the `likelihoodTools` package. 

We will show you how to create a model function, compute the MLE of the parameters 
(both steps from `likelihood` package), and then how to use the `likelihoodTools` 
package to: (i) extract the main information from the MLE results, (ii) plot 
residuals vs. predicted values, and (iii) plot the residuals vs. the observed values.

### Maximum Likelihood Estimation of the models parameters
As indicated in the `likelihood` package the first step is to formulate the model 
function, and then compute the MLE of the parameters. 


```{r load_data}
data(crown_rad)
dataset <- crown_rad
```


```{r model_fun}
modelfun <- function (a, b, DBH) {a + b * DBH}
```

Then, compute them MLE of the parameters using the `anneal` function from the 
`likelihood` package.

```{r mle}
results <- anneal(model = modelfun,
                  par = list(a = 0, b = 0),
                  var = list(DBH = "DBH", x = "Radius", mean = "predicted",
                             sd = 0.815585, log = TRUE),
                  source_data = dataset,
                  par_lo = list(a = 0, b = 0),
                  par_hi = list(a = 50, b = 50),
                  pdf = dnorm,
                  dep_var = "Radius",
                  max_iter = 20000,
                  show_display = FALSE)
```


### Extract the main information from the MLE results

The `likellihood::anneal()` function returns a list object with information on the annealing run. 
This list-type object contains several information (see `likellihood::anneal()` for more details).
The `mle_format()` function from `likelihoodTools` package allows extract the main 
information from the MLE results in a more user-friendly way. 

```{r mle_format}
mle_format(results, yvar = "Radius")
```

You can plot the table with more style using the`kableExtra` [@kableExtra] package.

```{r custom_table}
m <- mle_format(results, yvar = "Radius")

# Rename the columns
colnames(m) <- c(
  "Maximum Likelihood", 
  "# Params.", 
  "AIC~cor~", 
  "AIC", 
  "R^2^", 
  "Slope", 
  "RMSE"
)

# Print the results in a table
library(kableExtra)
kable(m, escape = FALSE, caption = "MLE Results", digits = 3) |> 
  kable_styling(full_width = F, position = "center", bootstrap_options = c("striped", "hover"))
```


### Plot the residuals vs. predicted values 

The `mle_plot_residuals()` function allows you to plot the residuals 
(observed - residuals) values *vs.* predicted values. The predicted 
values are obtained from the model with the parameters values estimated by maximum
likelihood estimation using simulated annealing.

```{r plot_residuals}
p_residuals <- mle_plot_residuals(results, yvar = "Radius")
```

### Plot the predicted vs. observed values

The `mle_plot_observed()` function allows you to plot the predicted values *vs.* 
observed values. The predicted values are obtained from the model with the parameters 
values estimated by maximum likelihood estimation using simulated annealing.

```{r plot_observed}
p_observed <- mle_plot_observed(results, yvar = "Radius")
```

### Customize the plot
If you don't like the default plot, you can customize it using  `ggplot2` [@@ggplot2] parameters.
For example, you can increase the size of the points, change the color, change axis titles, etc.

```{r plot_observed_custom}
p_observed +
  ggplot2::geom_point(size = 2, color = "red") +
  ggplot2::labs(x = "Observed Radius", y = "Predicted Radius") +
  ggplot2::theme(
    axis.title.x = ggplot2::element_text(size = 13, face = "bold"),
    axis.title.y = ggplot2::element_text(size = 13, face = "bold")
  )
```


### Plot the distribution of the residuals 

Maybe the user could be interested in plot the distribution of the residuals. 
We provided a code to do this. 

```{r residuals_dist}
library(ggplot2)
library(dplyr)

# compute the residuals
d <- results$source_data |> 
  dplyr::mutate(residuals = Radius - predicted)

# plot the histogram
p_residuals_dist <- ggplot(data = d, aes(x = residuals)) +
  geom_histogram(aes(y = after_stat(density), x = residuals), 
                   fill = "gray", col = "white", bins = 30) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Residuals") +
  ylab("Density")

p_residuals_dist 
```

Now, we can add the distribution to the histogram. For those, we used the `MASS` package
[@MASS]. 

```{r residuals_dist_fit}
library(MASS)

# estimate the distribution of the residuals  
dist <- MASS::fitdistr(d$residuals, "normal") 
params <- as.data.frame(t(dist$estimate))


p_residuals_dist <- p_residuals_dist  + stat_function(fun = dnorm, 
              args = list(mean = params$mean, 
                          sd = params$sd), 
              color = "black", size = 1) 
```


### More customization

Maybe the user is intereseted in arrange the residuals plots, the observed vs. 
predicted and the error distribution plots in a single plot. There are several packages 
as `patchwork` [@patchwork] or `cowplot` [@cowplot] that allow you to do this. Here, 
we provide a simple code to do this. 

```{r arrange_plots}
#| fig.width: 10
#| fig.height: 5

library(patchwork)


(p_observed + p_residuals + p_residuals_dist +  plot_annotation(tag_levels = "a")) & 
  theme(axis.title = element_text(size = 14, face = "bold"), 
        axis.text = element_text(size = 11),
        plot.tag = element_text(size = 14, face = "bold"))

```

# References

